---
title: "Regression and model validation"
author: "Marc Denojean-Mairet"
output: html_document
---



For this analysis, a data frame named **learningAnalysis2014** is created.  The CSV file named "learning2014.csv" is read.  The data frame consist of 7 variables (gender ,Age ,attitude, deep,  stra, surf, Points) and 166 observations.  The data are from a survey of statistics students.  The data include the global attitude of the students toward statistics and their exam points.  deep", "stra" and "surf" are combined variables by taking the mean.  "attitude" was scaled based on Likert scale (1-5) by dividing the "Attitude" column by 10.

More information about the data can be found **here** (https://www.mv.helsinki.fi/home/kvehkala/JYTmooc/JYTOPKYS3-meta.txt)
```{r}
learningAnalysis2014 <- read.csv(file = 'data/learning2014.csv')
dim(learningAnalysis2014)
str(learningAnalysis2014)
```
This is a plot of all the relationship among the variables.

```{r}

pairs(learningAnalysis2014[-1], col = "red")

library(ggplot2)
library(GGally)

ggpairs(learningAnalysis2014, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

ggplot(learningAnalysis2014, aes(x = attitude, y = Points, col = gender))+geom_point()+geom_smooth(method = "lm")+ggtitle("Student's attitude versus exam points")

summary(learningAnalysis2014)
```
