---
title: "Chapter 5"
author: "Marc Denojean-Mairet"
date: "04/12/2021"
output: html_document
---

# Chapter 5: Dimensionality reduction techniques

## Part 1: Show a graphical overview of the data and show summaries of the variables in the data

```{r message = FALSE, warning = FALSE}

# set plots size
knitr::opts_chunk$set(fig.width=16, fig.height=10) 

# uploading dataset from: http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt
human <- read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep = ",",  header = TRUE)

# code is from DataCamp.
# library
library(GGally)
library(corrplot)

# visualize the 'human_' variables
ggpairs(human)

# compute the correlation matrix and visualize it with corrplot
cor(human) %>% corrplot

# summary of the data
summary(human)
```
### Interpretation:

From the above graphs, we see that we have normal, binomial distributions.  We also see some strong negative and positve correlations.  For instance, Ado.Birth as a positive correlation with Mat.Mor (0.759).  As well, Mat.Mor and Life.Exp have a negative correlation (-0.857).  Another interesting correlation is the positive correlation between Life.Exp and Edu.Exp (0.789).  There is a positive correlation between GNI and Life.Exp (0.627) but it is lower than I would have expected.  Life.Exp does increase in countries with high GNI but not much.  There is a negative correlation between Ado.Birth and Edu.EXp.  This would suggest that the two variables are moving in opposite direction.

## Part 2: Perform principal component analysis (PCA) on the not standardized human data.

```{r message = FALSE, warning = FALSE}
# code is from DataCamp

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("red", "blue"))
```
 
## Part 3: Standardize the variables in the human data and repeat the above analysis. 

```{r message = FALSE, warning = FALSE}
# code is from DataCamp

# standardize the variables
human_std <- scale(human)

# print out summaries of the standardized variables
summary(human_std)

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_std)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("red", "blue"))

```

```{r message = FALSE, warning = FALSE}
# create and print out a summary of pca_human
s <- summary(pca_human)
s

# rounded percentages of variance captured by each PC
pca_pr <- round(100*s$importance[2,], digits = 1) 

# print out the percentages of variance
pca_pr

# create object pc_lab to be used as axis labels
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.8, 1), col = c("red", "blue"), xlab = pc_lab[1], ylab = pc_lab[2])
```

### Interpretation:

From the above tables and graphs, we see that the non-standardized data are difficult to interpret.  The biplot graph shows most of the country to be on the top right corner and the arrow to have a 0 length.  The standardized PCA biplot shows that most of the countries are in the middle of the graph.  We see that arrows of Mat.Mor and Ado.Birth have a small angle (meaning that they are correlated).  They have a positive correlation and are more explained by PC1.  GNI and Edu2.FM are strongly correlated and are explained by PC2.  Overall, the angles between arrows show correlations.  The smaller the angle the stronger the correlation.

## Part 4: personal interpretations of the first two principal component dimensions based on the biplot drawn after PCA on the standardized human data

### Interpretation:

PC1 captures 53.6% of total variance in the original variables.  PC2 captures 16.2% of variability.  We see that Ado.Birth, Life.Exp, Edue.FM, Mat.Mor, and GNI have a small angle with PC1.  This means that these variables have a high positive correlation with PC1.  The same is true with PC2 and Labo.FM and Parli.F. The length of the arrows are proportional to the standard deviations of the features.  We see that Mat.Mor, Labo.FM, and Life.Exp have the longest arrows.  Variables pointing in the direction of PC1 are contributing to that PC1 dimension.  The opposite is true for PC2.

## Part 5: Load the tea dataset from the package Factominer

```{r message = FALSE, warning = FALSE}
# code is from DataCamp

library(dplyr) 
library(ggplot2)
library(tidyr)
library(FactoMineR)

data("tea")

dim(tea)
str(tea)

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))

# look at the summaries and structure of the data
summary(tea_time)
str(tea_time)

# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)

# summary of the model
summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali", graph.type = "classic")
```

### Interpretation:

The Tea dataset relates to the consumption of tea and how, when, and where it is consumed.  There are 300 observations and 36 variables.  We see that Earl Grey is the favority type of tea and that it is mainly consumed alone.  The tea is mainly bought in chain store and consumed before or after lunch (not much during lunch).  The tea is from tea bags and there is an equal use or not of sugar with it.  We see that milk, Earl Grey, sugar, tea bags, and chain store are in the direction of Dim2.  This group is in the middle left side of the plot.  The other features are more dispersed and are in the right side of the plot.  Since these features are closer to the origin, we can say that more people use them.  For instance, Earl Grey vs black tea.  Early Grey is closer to the origin and therefore, more people use Early Grey than black tea.  This is more so with milk and lemon.