---
title: "Chapter 4: Clustering and classification"
author: "Marc Denojean-Mairet"
date: "27/11/2021"
output: html_document
---

# Chapter 4: Clustering and Classification

## Part 2: explore the structure and the dimensions of the data

Loading the Boston data.


```{r message = FALSE, warning = FALSE}

# set plots size
knitr::opts_chunk$set(fig.width=16, fig.height=10) 


#code from DataCamp.

#access the MASS package
library (dplyr)
library(MASS)
library(corrplot)
library(tidyr)


#load the data
data("Boston")

#explore the dataset
str(Boston)
dim(Boston)
summary(Boston)
```
## Description:
The Boston data set represent the housing values in Suburbs of Boston.
<br><br>
The Boston data frame has **506** rows and **14** columns.
<br><br>
This data frame contains the following columns:
<br><br>
**crim** = per capita crime rate by town<br>
**zn** = proportion of residential land zoned for lots over 25,000 sq.ft<br>
**indus** = proportion of non-retail business acres per town<br>
**chas** = Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)<br>
**nox** = nitrogen oxides concentration (parts per 10 million)<br>
**rm** = average number of rooms per dwelling<br>
**age** = proportion of owner-occupied units built prior to 1940<br>
**dis** = weighted mean of distances to five Boston employment centres<br>
**rad** = index of accessibility to radial highways<br>
**tax** = full-value property-tax rate per $10,000<br>
**ptratio** = pupil-teacher ratio by town<br>
**black** = 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town<br>
**lstat** = lower status of the population (percent)<br>
**medv** = median value of owner-occupied homes in \$1000s<br>

## Source:

Harrison, D. and Rubinfeld, D.L. (1978) Hedonic prices and the demand for clean air. J. Environ. Economics and Management 5, 81–102.

Belsley D.A., Kuh, E. and Welsch, R.E. (1980) Regression Diagnostics. Identifying Influential Data and Sources of Collinearity. New York: Wiley. 

## Part 3: graphical overview of the data 

```{r message = FALSE, warning = FALSE}

#plot matrix of the variables
pairs(Boston, gap=1/30)

# MASS, corrplot, tidyr and Boston dataset are available

# calculate the correlation matrix and round it
cor_matrix<-cor(Boston) %>% round(digits = 2)

# print the correlation matrix
cor_matrix

# visualize the correlation matrix
corrplot(cor_matrix, method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex = 0.6)

corrplot(cor_matrix, method = 'number', type="upper")

```

```{r message = FALSE, warning = FALSE}
# center and standardize variables
boston_scaled <- scale(Boston)

# summaries of the scaled variables
summary(boston_scaled)

# class of the boston_scaled object
class(boston_scaled)

# change the object to data frame
boston_scaled <- as.data.frame(boston_scaled)
```

```{r message = FALSE, warning = FALSE}
# summary of the scaled crime rate
summary(boston_scaled$crim)

# create a quantile vector of crim and print it
bins <- quantile(boston_scaled$crim)
bins

# create a categorical variable 'crime'
crime <- cut(boston_scaled$crim, breaks = bins, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))

# look at the table of the new factor crime
table(crime)

# remove original crim from the dataset
boston_scaled <- dplyr::select(boston_scaled, -crim)

# add the new categorical value to scaled data
boston_scaled <- data.frame(boston_scaled, crime)

# number of rows in the Boston dataset 
n <- nrow(boston_scaled)

# choose randomly 80% of the rows
ind <- sample(n,  size = n * 0.8)

# create train set
train <- boston_scaled[ind,]

# create test set 
test <- boston_scaled[-ind,]

# save the correct classes from test data
correct_classes <- test$crime

# remove the crime variable from test data
test <- dplyr::select(test, -crime)
```

```{r message = FALSE, warning = FALSE}
# linear discriminant analysis
lda.fit <- lda(crime ~ ., data = train)

# print the lda.fit object
lda.fit

# the function for lda biplot arrows
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "orange", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}

# target classes as numeric
classes <- as.numeric(train$crime)

# plot the lda results
plot(lda.fit, dimen = 2, col = classes, pch = classes)
lda.arrows(lda.fit, myscale = 1)

```


```{r message = FALSE, warning = FALSE}
# predict classes with test data
lda.pred <- predict(lda.fit, newdata = test)

# cross tabulate the results
table(correct = correct_classes, predicted = lda.pred$class)
```

```{r message = FALSE, warning = FALSE}
# load MASS and Boston
library(MASS)
data('Boston')

# euclidean distance matrix
dist_eu <- dist(Boston)

# look at the summary of the distances
summary(dist_eu)

# manhattan distance matrix
dist_man <- dist(Boston, method = 'manhattan')

# look at the summary of the distances
summary(dist_man)
```

```{r message = FALSE, warning = FALSE}
# k-means clustering
km <-kmeans(Boston, centers = 3)

# plot the Boston dataset with clusters
pairs(Boston, col = km$cluster)
```

```{r message = FALSE, warning = FALSE}
model_predictors <- dplyr::select(train, -crime)

# check the dimensions
dim(model_predictors)
dim(lda.fit$scaling)

# matrix multiplication
matrix_product <- as.matrix(model_predictors) %*% lda.fit$scaling
matrix_product <- as.data.frame(matrix_product)


#install.packages("plotly")
library(plotly)
plot_ly(x = matrix_product$LD1, y = matrix_product$LD2, z = matrix_product$LD3, type= 'scatter3d', mode='markers')
```

