---
title: "Chapter#3: Logistic regression"
author: "Marc Denojean-Mairet"
date: "18/11/2021"
output: html_document
---

# Logistic regression

Read new data frame named alc.csv to studentAlc and print variable names.


```{r}


#studentAlc <- read.table("~/IODS-project/IODS-project/data/pormath.csv", sep = ",", header = TRUE)

studentAlc <- read.table("~/IODS-project/IODS-project/data/alc.csv", sep = ",", header = TRUE)

dim(studentAlc)

colnames(studentAlc)

```

**Data Set Information:**

"This data approach student achievement in secondary education of two Portuguese schools. The data attributes include student grades, demographic, social and school related features) and it was collected by using school reports and questionnaires. Two datasets are provided regarding the performance in two distinct subjects: Mathematics (mat) and Portuguese language (por). In [Cortez and Silva, 2008], the two datasets were modeled under binary/five-level classification and regression tasks. Important note: the target attribute G3 has a strong correlation with attributes G2 and G1. This occurs because G3 is the final year grade (issued at the 3rd period), while G1 and G2 correspond to the 1st and 2nd period grades. It is more difficult to predict G3 without G2 and G1, but such prediction is much more useful (see paper source for more details)."

The above information is from UCI Machine Learning Repository.

More information about the data sets can be found here:(https://archive.ics.uci.edu/ml/datasets/Student+Performance)


I chose the following 4 variables to predict high use of alcohol:  G3, absences, Pstatus, and health.  I have chosen these variables thinking that they would be easily available to a school without the need to survey students.

**G3**:

My assumption here is that low grade is an indication of high alcohol use since high use of alcohol can affect cognitive functions such as memory.

**absences**:

High number of absences could be an indication of high alcohol use as it would affect one's schedule.  Absenses could be due to being sick after consuming a lot of alcohol.

**goout**

Going out with friend a lot might raise alcohol consumption since there will be more opportunities to consume alcohol.


**health**:

Poor health could be an indication of high alcohol consumption.  Alcohol can negatively affect physical and mental health.

### These are plots of all variables

```{r}
# access the tidyverse libraries tidyr, dplyr, ggplot2
library(tidyr); library(dplyr); library(ggplot2)
# draw a bar plot of each variable
gather(studentAlc) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()

```

### Box plots of my chosen variables vs high-use of alcohol

```{r}
# initialize a plot of high_use and G3
g1 <- ggplot(studentAlc, aes(x = high_use, y = G3, col = sex))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("grade") + ggtitle("Student final grade by alcohol consumption and sex")

# initialise a plot of high_use and absences
g2 <- ggplot(studentAlc, aes(x = high_use, y = absences, col = sex))

# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ggtitle("Student absences by alcohol consumption and sex")

# initialize a plot of high_use and 
g3 <- ggplot(studentAlc, aes(x = high_use, y = goout, col = sex))

# define the plot as a boxplot and draw it
g3 + geom_boxplot() + ylab("going out") + ggtitle("Student going out with friends by alcohol consumption and sex")

g4 <- ggplot(studentAlc, aes(x = high_use, y = health, col = sex))

# define the plot as a boxplot and draw it
g4 + geom_boxplot() + ylab("health") + ggtitle("Student health by alcohol consumption and sex")
```

## Numerical exploration of my chosen variables

```{r}
# produce summary statistics by grade
studentAlc %>% group_by(sex, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

# produce summary statistics by absences
studentAlc %>% group_by(sex, high_use) %>% summarise(count = n(), mean_absences = mean(absences))

# produce summary statistics by going out
studentAlc %>% group_by(sex, high_use) %>% summarise(count = n(), mean_going_out = mean(goout))

# produce summary statistics by health
studentAlc %>% group_by(sex, high_use) %>% summarise(count = n(), mean_health = mean(health))
```

## Interpretation of the above plots and tables

We see that grades are lower for male when high_use of alcohol is true.  Female seems to be less negatively affected.  Final grade is not as a strong predictor as I thought.  Absences and high_use of alcohol seems to have a good correlation.  Female seems to have a little more absences than male when high_use of alcohol is true.  going out shows some correlation for male and female for high_use of alcohol.  The more a student go out the more he or she consume alcohol.  Health does not seem to have a strong correlation with high_use of alcohol.  I would have thought that there would have been a stronger correlation.  It is possible that the participants did not truthfully answer this question or that participants are not aware of their overall health (mental and physical).  This seems to be especially true for male.  Female have a broader range of answers with 25% of them describing their health below 2 (Q1 high_use = true).

## Logistic regression to statistically explore the relationship between my chosen variables


For this model, high_use is the target variable and final grade, absences, going out, and health are the predictors.
I did not include sex because if a school needs to predict high_use of alcohol, sex might add an unnecessary bias.  For instance, male students might be watched more carefully than female students because male seems to have higher high_use of alcohol.
```{r}
# find the model with glm()
m <- glm(high_use ~ G3 + absences + goout + health, data = studentAlc, family = "binomial")

# print out a summary of the model
summary(m)

# print out the coefficients of the model
coef(m)

# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI <- confint(m) %>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```

## Interpretation of the above model

From the model P-values, we see that final absences and going out (goout) are likely relevant variables to explain high_use of alcohol.  However, final grade and health show that the data are providing little evidence that these variables are needed to explain high_use.

From the Odd Ratio table, we see that absences, going out, and health have an OR greater than 1.  This would imply that these variables are positively associated with high_use of alcohol.  Going out has a Odd Ratio of 2 showing a very strong positive association with high_use of alcohol.  Final grade is close to 1 as well.  This would imply that the positive association is not as strong as the other variables.

It seems that most of my chosen variables have some positive association with high_use.  Therefore, they could be used to correctly predict high_use.

## Predictive power of my model

```{r}
# predict() the probability of high_use
probabilities <- predict(m, type = "response")

# add the predicted probabilities to 'studentAlc'
studentAlc <- mutate(studentAlc, probability = probabilities)

# use the probabilities to make a prediction of high_use
studentAlc <- mutate(studentAlc, prediction = probability > 0.5)

# see the last ten original classes, predicted probabilities, and class predictions
select(studentAlc, G3, absences, goout, health, high_use, probability, prediction) %>% tail(10)

# tabulate the target variable versus the predictions
table(high_use = studentAlc$high_use, prediction = studentAlc$prediction)

# initialize a plot of 'high_use' versus 'probability' in 'studentAlc'
g <- ggplot(studentAlc, aes(x = probability, y = high_use, col = prediction))

# define the geom as points and draw the plot
g + geom_point()

# tabulate the target variable versus the predictions
table(high_use = studentAlc$high_use, prediction = studentAlc$prediction) %>% prop.table %>% addmargins

# define a loss function (average prediction error)
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

# call loss_func to compute the average number of wrong predictions in the (training) data
loss_func(class = studentAlc$high_use, prob = studentAlc$probability)


```

