---
title: "Regression and model validation"
author: "Marc Denojean-Mairet"
output: html_document
---



For this analysis, a data frame named **learningAnalysis2014** is created.  The CSV file named "learning2014.csv" is read.  The data frame consist of 7 variables (gender ,Age ,attitude, deep,  stra, surf, Points) and 166 observations.  The data are from a survey of statistics students.  The data include the global attitude of the students toward statistics and their exam points.  deep", "stra" and "surf" are combined variables by taking the mean.  "attitude" was scaled based on Likert scale (1-5) by dividing the "Attitude" column by 10.

More information about the data can be found **here** (https://www.mv.helsinki.fi/home/kvehkala/JYTmooc/JYTOPKYS3-meta.txt)
```{r}
learningAnalysis2014 <- read.csv(file = 'data/learning2014.csv')
dim(learningAnalysis2014)
str(learningAnalysis2014)
```
These are plots of all the relationship among the variables.  From the visualization, we see some positive and negative correlations.  An interesting correlation is attitude and points.  As expected, we see a negative correlation with surf and deep.  The most negative correlation is with deep and points.  We also see that there are more female than male students.  However, from the plots, there is no good fit between gender and points and there is no strong correlations.

The summary table gives us more information about the means of the variables.  

```{r}

pairs(learningAnalysis2014[-1], col = "red")

library(GGally)
ggpairs(learningAnalysis2014, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

summary(learningAnalysis2014)


```


```{r}
ggpairs(learningAnalysis2014, lower = list(combo = wrap("facethist", bins = 20)))

#my_model <- lm(points ~ attitude + stra + surf, data = learningAnalysis2014)

#summary(my_model)

```

