---
title: "Chapter 6"
author: "Marc Denojean-Mairet"
date: "10/12/2021"
output: html_document
---

# Chapter 6: Analysis of longitudinal data

## Part 1: Implement the analyses of Chapter 8 of MABS

### RATS dataset
```{r message = FALSE, warning = FALSE}
# most of the code is from DataCamp

# set plots size
knitr::opts_chunk$set(fig.width=16, fig.height=10)

# Loading RATS dataset

# libraries
library(dplyr)
library(tidyr)

# load rats data
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep = "\t", header = TRUE)

# Factor variables ID and Group
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

RATS

# Convert data to long form
RATSL <- RATS %>%
  gather(key = WD, value = Weight, -ID, -Group) %>%
  mutate(Day = as.integer(substr(WD,3,4))) 

# dataset
print(RATSL, row.names = FALSE)

```

### Plot of the RATSL data

```{r message = FALSE, warning = FALSE}

#Access the package ggplot2
library(ggplot2)

# Glimpse the data
glimpse(RATSL)

# Draw the plot
ggplot(RATSL, aes(x = Day, y = Weight, linetype = ID, color = ID)) +
  geom_line() + geom_point() +
  facet_grid(. ~ Group, labeller = label_both)

```


From the above plots, we see that each group of rats started with a different initial weight.  Group 1 gained the least amount of weight during the study period.  Group 2 and group 3 did gain weight at approximately the same rate.

### Plots of the RATSL data after standartization

```{r message = FALSE, warning = FALSE}

# Standardise the variable RATS
RATSL <- RATSL %>%
  group_by(Day) %>%
  mutate(stdweight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL)

# Plot again with the standardised RATS
ggplot(RATSL, aes(x = Day, y = stdweight, linetype = ID, color = ID)) +
  geom_line() + geom_point() +
  scale_linetype_manual(values = rep(1:16, Days=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized Weight")

```

From the above standardized plots, we see that group 1 did not gain or lose weight.  We also see that group 2 gained the most weight and group 3 lost weight.

### Summary Graph

```{r message = FALSE, warning = FALSE}

# Number of weeks, baseline (week 0) included
n <- 9

# Summary data with mean and standard error of RATS by Group and Day
RATSS <- RATSL %>%
  group_by(Group, Day) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSS)

# Plot the mean profiles
ggplot(RATSS, aes(x = Day, y = mean, linetype = Group, shape = Group, color = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

```

From the summary plot with mean and standard error, we see that group 1 did not gain much weight during the study but group 2 and group 3 did gain weight at a similar rate.


### Outliners

```{r message = FALSE, warning = FALSE}

# Create a summary data
RATSL8S <- RATSL %>%
  filter(Day > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL8S)

# Draw a boxplot of the mean versus treatment
ggplot(RATSL8S, aes(x = Group, y = mean, color = Group)) +
  geom_boxplot(outlier.colour="blue", outlier.shape=16,
             outlier.size=10, notch=FALSE) +
  stat_summary(fun = mean, geom = "point", shape=23, size=10, fill =  "red", colour = "blue") +
  scale_y_continuous(name = "mean(Weight), weeks 1-9")


```

From the above boxplot, group 2 seems to have more variation than group 1 and 3.  As well, we see 3 outliers.


```{r message = FALSE, warning = FALSE}
# Add the baseline from the original data as a new variable to the summary data
RATSL8S2 <- RATSL8S %>%
  mutate(baseline = RATS$WD1)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline + Group, data = RATSL8S2)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)



```

Given that Pr(>F) is smaller than 0.05, we reject the null hypothesis (all means are equal) and we conclude that at least one group is different in weight than the others.

## Part 2: Implement the analyses of Chapter 9 of MABS

```{r message = FALSE, warning = FALSE}

```

